% Version: 0.0

\documentclass[../main.tex]{subfiles}

\begin{document}
% General context
%% Put synergy, a term from information theory, at home in the computational science mindset
In complex systems, elementary properties of many individuals lead to emergent behavior on a system-size scale.
We can view this emergence as a ’synergy’ of system elements together, which is a vital part of the dynamics of the system as a whole.
In information theory, synergy as a quantifiable property of a system is a relatively new idea, and can be seen as an addition to system properties like entropy and mutual information.
The method of quantification for systems larger than the trivial case of 2 random variables is amongst the open problems in the field, but multiple measurements have been proposed \cite{olbrich2015information}.
Synergy and mutual information are closely related properties, as synergy can be seen as negative mutual information; it is additional information on a random variable that is given by a combination of other random variables, yet not a part of the information captured in any of the individual random variables.
As such, visualizations of the mutual information in a system of dependent random variables can sketch an image of the synergy in a system.
One method of visualization is through mutual information profiles.
Early versions used pairwise mutual information between all variables to create a profile of a system \cite{bar2013computationally}. 
The extended full mutual information profile, proposed by Quax et al., uses the mutual information between all possible subsets of input variables with an output variable \cite{quax2017quantifying}.
This is hypothesized that it would allow for the identification whether synergy is present at a low level, between two variables, or at a higher level, between groups of variables.

%% Link to gene regulatory networks/biology
%% Talk about pertubation resistance
Many natural systems are complex in their nature, such as neural networks, gene regulatory networks, bird flocking patterns, and pattern formation \cite{choi2001supply, gat1999synergy, kondo2010reaction, liang2008gene}.
A complex system is a system that showcases emergent behavior from the interaction of single participants in the system.
Complexity itself is referred to in two manners in the literature, in both in a qualitative manner of which the properties characterize complex systems, and as a quantification of the added information in a system that comes only from the interactions between variables \cite{bar2004multiscale}.
Synergy is a quantification of this added information that makes the whole more than the sum of its parts.
In this study, we refer to complexity in the latter sense.
Many of these complex natural systems are also complicated; they consist of many members, sometimes of different species, with many relationships of various kinds running inbetween.
An example is the food web of a tropical rainforest, where many different species participate and diets are so diverse that the network is densily connected.
These networks are relatively simple in their building blocks, but show complicated patterns on a global scale.
In gene regulation, the expression of genes is spatially regulated from the conception of the organism onwards.
A possible angle to the question 'why are biological complex systems often complicated' is from the perspective of resilience and memory.
Biological systems require a level of memory, a relation between the current state of the system and previous states of the system. 
They also require a level of noise resilience, as biological systems tend to experience shocks from external sources \cite{peixoto2012emergence}. % not the best reference
Maximizing one is not always in the best interest of the other; noise resilience is maximized when the system automatically defaults to a hard-coded state, but this leaves no room for system memory for any state but the hard-coded default.
A maximized system memory, on the other hand, will never forget noise, causing noise to never die out over time.
It has been hypothesized by Quax et al. that synergy increases the resilience of a system against pertubation in a single input variable \cite{quax2017quantifying}.
This means that synergy can be utilized to make a system resistant to nudges, while retaining the ability to memorize previous states.
We assume that the nudges that biological systems experience the most, and should be resilient against, are single-variable nudges.
If true, this would fit well with the proposal that synergy is used to provide protection against these disturbances.
After all, synergy operates at a multi-variate level, whereas single variable nudge operate at a single-variate level.
The realization of a middle way, that maximizes the combination of both resilience and memory, might be through synergy.
As such, synergistic relationships in the system stemming from complicatedness could contribute to its functioning in a noisy environment.

Thus far, the primary method of investigating complex natural networks is by looking at 1-to-1 relationships between variables \cite{}. % add several sources
For instance, to build and understanding of gene regulation we examing the correlation in the expression between pairs of genes.
Multivariate relationships remain unexplored in these studies. 
The potential presence of a relationship between synergy and resilience makes it very interesting to determine the amount of synergy in a biological network. 
This is even more so the case considering that the relationship between system complexity and resilience is one of the primary unanswered questions of ecology \cite{}.
The amount of synergy in biological complex systems has, to our best knowledge, not been quantified or explained.

% In this study, we want to...
In this work we aim to examine the links between the complicatedness of a complex system its resilience, and the memory of the system.
In particular, we are interested in the role of synergy in these networks, which is used in this study as a quantification of the system complexity.
This problem is recurrent in many other disciplines, as principles from information theory are broadly applied.
We hypothesize that there is a form of synergistic control in complex biological networks.
As a first step, we aim to quantify the amount of synergy in a biological complex system.
This should allow us to determine how big the presence of synergy is in real-world systems.
We focus on gene regulation networks, as small, elementary motifs are readily available in these networks.
Resilience is a hot topic in gene regulation, has been examined through discrete Boolean networks without taking synergy into account \cite{peixoto2012emergence}.
We expect that this discrete approximation is not always sufficient for large networks, as it has been found that the same Boolean motif does not always have the same function \cite{ingram2006network}.
Sticking to a small network size helps both in the quantification of synergy, as computation of the synergy is an expensive operation for large systems \cite{jointpdf}.
In addition to synergy, we also aim to quantify the system memory and resilience.
This will provide a starting frame of reference for both measures, and give an idea to what extent natural networks are resilient to perturbation and capable of remembering previous states.
Secondly, we test the hypothesis that a gene regulation network has more synergy than a random gene regulation network of a similar size. 
We investigate this through a simulation study in continuous space, where networks are approximated locally in time through an ODE system.
In addition, we also measure whether real-world motifs score better in terms of system memory and single-variate nudge resilience than random networks, and whether these networks are Pareto optimal in these two properties.
Thirdly, we want to test our assumption that biological networks should be resilient to single-variable nudges, but are not necessarily resilient to nudges in multiple variable at once.
We do so by examining the resilience of a real GRN when nudging an increasing number of variables.
Finally, we take a more detailed look at the level at which synergy occurs using full mutual information profiles.
With this, we provide insight into larger gene regulation motifs, as well as a case for the use of these profiles in the analysis of synergy in complex systems.
\end{document}